---
title: "Class 1"
author: "Florence"
date: "2025-10-16"
output: html_document
---

```{r}
 x <- c(0.5, 0.6)
 age<-c(20,35,32,29)
 summary(age)
```

```{r}
  x <- c(0.5, 0.6)
  age <- c(20,35,32,29)
  summary(age)
  age <- c(35,24,18,24)
  mean(age)
```

#Logical
```{r}
 x <- c(TRUE, FALSE)
 x <- c(T, F)
```

#Character
```{r}
 x<-c("a","b","c")
 class<-c("M","F","F","M")
```

# Integer
```{r}

 x<-9:29
```

#Complex
```{r}

 x<-c(1+0i,2+4i)
```

#Matrix and operation within matrices
```{r}
 x<-matrix(1:6,nrow=2,ncol=3)
 x
 (y<-matrix(1:6,nrow=3,ncol=2))
 
  x%*%y
```

#List
```{r}
x<-list(1,"a",TRUE,1+4i)
x
```
```{r}
 x<-factor(c("yes","yes","no","yes","no"))
```

#Create a vecto rwith NAs in it
```{r}
 x<-c(1,2,NA,10,3)
 is.na(x)
      
```


```{r}
 x<-c(1,2,4,"NA",5)
 bad<-is.na(x)
 print(bad)
 
 x[!bad]
```


```{r}
x<-c(1,2,NA,4,NA,5)
 y<-c("a","b",NA,"d",NA,"f")
 good<-complete.cases(x,y)
 good
 
 sum(c(TRUE,TRUE,FALSE,FALSE,FALSE))
```

#Creating data frame
```{r}
 studentID<-c(1,2,3,4,5)
 math_score<-c(12,17,10,9,NA)
 gender<-c("M","F","M","M","F")
 it_score<-c(13,18,11,10,19)
 scoredata<-data.frame(studentID,gender,math_score,it_score)
 scoredata

View(scoredata)
```

#Create dataset and read it from your computer

```{r}
data(data_class2)

data_class2<-data.frame(height=numeric(0),weight=numeric(0),bmi=numeric
 (0))
 data_class2<-edit(data_class2)
```

#read a dataset from computer
```{r}
data_class <- read.table("C:\\Users\\USER\\Documents\\R programming\\Datasets\\Book1.txt", header = TRUE)

variable.names(data_class)

```


```{r}
 head(data_class)
```


```{r}
tail(data_class)
```

```{r}
data_class[10:20,]
```

```{r}
summary(data_class$WEIGHT)
```


```{r}
length(data_class$WEIGHT)
```


```{r}
data_class[,-1]
```
#Use data set built in R
 
```{r}
data("airquality")
variable.names(airquality)
 str(airquality)
```

#Exploration data analysis
```{r}
data_class
```


```{r}
 data("iris")
 head(iris)
```

#mean(iris$Petal.Length)
```{r}
mean(iris$Petal.Length)
```
```{r}
mean(iris$Petal.Length[iris$Species=="setosa"])
```
```{r}
mean(iris$Petal.Length[iris$Species=="versicolor"])
```
```{r}
 mean(iris$Petal.Length[iris$Species=="virginica"])
```

#shortcut
```{r}
 by(iris$Petal.Length,iris$Species,mean)
```

```{r}
by(iris$Petal.Length,iris$Species,sd)
```

```{r}
by(iris$Petal.Length,iris$Species,summary)
```

#Visualize the differences in continuous variables between categories using Box-and whisker plot

```{r}
 boxplot(iris$Sepal.Length~iris$Species,data=iris, main="Comparison")
```

#Relationship between two categorical variables

```{r}
ucba<-data.frame(UCBAdmissions)
 head(ucba) 
 
```


```{r}
cross<-xtabs(Freq~Gender+Admit,data=ucba)
 (cross<-xtabs(Freq~Gender+Admit,data=ucba))
```
#Is there gender bias in UCBgraduate admission process?

```{r}
 prop.table(cross,2)
```
Simpson’s paradox
```{r}
cross2<-xtabs(Freq~Gender+Admit,data=ucba[ucba$Dept=="A",])
 prop.table(cross2,1)
```

```{r}
dat <- read.table(text = "
ProdA ProdB ProdC ProdD ProdE
1 110 50 60 70
2 120 50 80 65
", header = TRUE)


 barplot(as.matrix(dat),beside=FALSE,col=c("Red","green"))
```


```{r}
#barplot(as.matrix(dat),beside=TRUE,col=c("gold3","red"))

 dat <- read.table(text = "A B C D E F G
 1 10 80 30 90 70 60 90
 2 20 50 70 50 40 10 40
 3 60 80 80 60 60 30 160
 4 20 40 70 80 20 10 70", header = TRUE)
 
 barplot(as.matrix(dat))
```
#Scatter Plot
 
```{r}
plot(WEIGHT~HEIGHT,data=data_class) 


```


```{r}
plot(data_class$HEIGHT,data_class$WEIGHT)
```

```{r}
 plot(data_class[,1],data_class[,2])
```
 
 
```{r}
cor(data_class$HEIGHT,data_class$WEIGHT)

```

#Linear Models
```{r}
plot(WEIGHT~HEIGHT,data=data_class)
abline(lm(WEIGHT~HEIGHT,data=data_class)$coefficient) 
 
```

```{r}
data_lm<-lm(WEIGHT~HEIGHT,data=data_class)

summary(data_lm) 

```


```{r}
hist(data_class$HEIGHT)
```


```{r}
 hist(data_class$WEIGHT,col = "blue")
```
#Linear with other line
```{r}
 plot(WEIGHT~HEIGHT,data=data_class)

 abline(lm(WEIGHT~HEIGHT,data=data_class)$coefficient)
 
 lines(lowess(data_class$HEIGHT,data_class$WEIGHT),col="blue") 
```

```{r}
cor(data_class$HEIGHT,data_class$WEIGHT)
```

```{r}
 model1<-lm(WEIGHT~HEIGHT,data=data_class)

```

#Data management

#Create another Variable
 
```{r}
attach(data_class)

(BMI<-WEIGHT/(HEIGHT/100)^2)
```


```{r}
(BMI<-round(WEIGHT/(HEIGHT/100)^2,digit=2))
```


```{r}
head(cbind(data_class,BMI),n=5)
```


```{r}
new_data_class<-cbind(data_class,BMI)
 tail(cbind(data_class,BMI),n=10)
```

#Summary of BMI
```{r}
 summary(BMI)
```


```{r}
stem(BMI,scale=2)
```
 
 
